---
title: "Repression Strategies and Mass Mobilization"
author:
- affiliation: University of Iowa
  name: Yufan Yang
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  pdf_document:
    citation_package: natbib
    fig_caption: yes
    keep_tex: yes
    latex_engine: pdflatex
    template: svm-latex-ms2.tex
citecolor: black
biblio-style: apsr
endnote: no
fontsize: 11pt
keywords: repression, mobilization
link-citations: yes
linkcolor: black
bibliography: \dummy{`r file.path(getwd(), list.files(getwd(), "bib$"))`}
spacing: double
abstract: 'Previous studies have provided competing theories and mixed empirical evidence
  on how repression effects mass mobilization. I argue that the answer to this question
  lies in different repression strategies. Especially, physical rights regression
  such as political killing, imprisonment without a legal process, and torture, impairs
  regime legitimacy and provides willingness for the public to participate in collective
  action. It also provides opportunities for dissent groups to expand. Hence, physical
  rights repression tends to escalate mass mobilization instead of deterring it. In
  contrast, censorship, which is the block of information, can isolate the country
  from the outside world, isolate the dissent group from the public, and protect the
  government’s credibility. Therefore, censorship lowers the willingness for the public
  to mobilize and reduces the opportunities for dissent groups to expand. Hence, censorship
  can successfully deter mobilization. '
---
\section{Introduction}

The interaction between repression and mobilization is a long-standing debate. Political scientists have agreed that mass mobilization leads to higher repression, but there is still no consensus on whether repression successfully deters or escalates mass mobilization. Studies of how effective state repression is have provided competing theories and mixed evidence. Different arguments include: an inverted-U relationship between state repression and conflict [@Muller1990], a positive impact that repression escalates conflict [@MasonandKrane1989; @Goodwin2001; @Hultquist2017], and a negative impact that repression deters or defeats conflict [@Downes2008, @Lyall2009]. Further, some studies explore the dynamics between repression and mobilization by looking at different levels or different types of repression. For example, @Lichbach1987 builds a rational actor model and finds that low-level repression tends to deter violent collective action when fear replaces anger. However, @Gupta1993 argue that high-level repression can also be effective in autocracies. Nonetheless, Moore’s [-@Moore1998] empirical evidence supports Lichbach’s [@Lichbach1987] theory instead of Gupta’s [-@Gupta1993]. Later, @Carey2006 claims that Moore’s [-@Moore1998] analysis is flawed because he did not take account of time and argues that repression is always ineffective in democracies. 

In fact, the answer to this question may lie in different types of repression, as some scholars have noticed, that violent and non-violent repression probably has distinct impact on mobilization behaviors. Violent repression refers to physical torture, killing, disappearing, etc., while non-violent repression usually defined as constraints on civil rights, especially civil rights of mobilization. Nonetheless, the development of technology, especially computer science, provides another strong tool for the government to repress: censorship. Censorship, as an effective way to block information, widely exists in autocracies. The public can only receive the information that the state allows them to receive, contributing to regime legitimacy and social stability. For instance, one of the reasons that repression escalates mass mobilization is that, state violence can damage regime legitimacy, so that the willingness for the public to initiate anti-government collective action as well as the opportunities for dissent groups to expand increase @Greene1974. However, if the state successfully hides the information on state violence, the public would not have willingness for mass mobilization. Further, civil rights as a western, democratic term, can also be manipulated by state censorship in autocracies, making the public accept the state’s ideology and lowering the willingness for collective action. That is to say, a combination of conventional repression and censorship can be an effective way to deter mass mobilization. 
    
Therefore, it is necessary to distinguish these three types of repression and explore the dynamics between each one and mass mobilization. This paper builds a model using differential equations to identify the dynamic interactions within a system of physical rights repression, civil rights repression, censorship, and mass mobilization. This paper applies a dynamic model using differential equations to explore how repression and mass mobilization interact with each other. More specifically, this paper categories repression strategies into three categorizes: physical rights repression, civil rights repression, and censorship. This paper finds that, if a state mostly relies on violent repression, mass mobilization tends to escalate, which is a backfire effect. However, when the state is aware of the growth of mass mobilization, it will intensify censorship to isolate the dissent group. With intensified censorship and increased violent repression, mass mobilization can be successfully deterred. Further, if a state has a high level of censorship at the beginning, the level of mobilization tends to be stable in the short term, which is unlikely to be a threat to state authority. After a certain time, mass mobilization can still slightly increase, but physical rights repression and censorship will grow correspondingly. After censorship and physical rights repression reaches the peak, mass mobilization can be successfully deterred as well.

\vspace{3mm}

\section{Repressive Strategies and Mass Mobilization}

States have various tools of repression. On the one hand, the state can use violence to directly eliminate the dissents’ capability of mobilization. These tools include: killings by the government without due process of law; making people disappeared out of political reasons; incarcerating citizens out of political reasons; and tortures of individuals from the government. These tools are aimed to make dissents physically unable to make any collective action, and they are categorized as physical rights repression in this paper. 

On the other hand, the state can limit the dissents’ capability of mobilization in two ways. First, the state can constraint civil rights such as freedom of movement, freedom of religion, freedom of ideology, etc., which reduces the opportunities of mass mobilization. These tools are categorized as civil rights repression in this paper. Second, with the help of technology, the state can simply block information that threatens regime legitimacy or state authority. This information can be news that reveals states’ violation of human rights, news that reveals government officials such as corruption, as well as ideologies of democracy, liberty, human rights, and so on. Blocking threatening information can lower the public’s willingness of mobilization, and thus is widely applied in autocracies. Physical rights repression, civil rights repression, and censorship can have different even opposite impact on mobilization, and this is why previous studies have presented distinct empirical results. 

\section{Physical Rights Repression}

Physical rights repression is violent. It might be the most effective way to stop individuals from joining collective action, but it tends to be counterproductive overall. For one thing, state violence against civilians impairs the regime legitimacy and provides more willingness for the dissent group and the public to escalate mass mobilization (@Goodwin2001). For another, state violence, especially indiscriminate violence against civilians, may even convince the public who are previously inactive to join the dissent group to seek for protection and survival (@Kalyvas2007). In short, physical rights repression increases willingness for the dissents and general public of collective action, as well as opportunities for the dissents to expand their influence. Therefore, physical rights repression alone is at best ineffective and at worst counterproductive (@Kalyvas2006). 

\textit{Hypothesis 1: Physical rights repression escalates mass mobilization.}

\section{Civil Rights Repression}

Civil rights repression can be complicated. To reduce the dissents’ capability of mobilization without using violence, the state tends to constrain freedom of peaceful assembly, freedom of religion, freedom of speech, and freedom of movement. These tools are usually pre-emptive and can be effective if the state use these tools constantly. Though the violation of these human rights can lead to social frustration and anger, constant repression on civil rights can keep the opportunities of mass mobilization at a low level. Nonetheless, constant repression on civil rights requires considerable investment and detailed implementation strategies, such as governmental spending on police and secret police in order to monitor and control civilians. Governmental investment on civil rights repression can fluctuate over time, and given this inconsistency, it is hard to tell whether civil rights repression can effectively deter mass mobilization in reality. 

\section{Censorship}

Hiding, blocking, and manipulating information is the subtlest way of repression, and sometimes the public are not even aware of this type of repression since the information is not available. As previously mentioned, two types of information are the most likely to be censored: negative information on the government and information of democracy, liberty, human rights, and collective action (@King2013). Censorship tends to be the most effective repression strategy for the following reasons. First, censorship isolates the country from the outside world, and thereby democratic values are not easily to be passed on to citizens in authoritarian regimes. Even if the public is aware of democratic values, it is unlikely that they are fully convinced by them, which happens in China (@LuandShi2015), because the state has the power to decide what kind of information its citizens can receive. Therefore, the willingness of the public to start anti-regime mobilization towards autocracies is low. Second, censorship helps to hide negative news about the government and government officials. The state may censor the information that would damage the image, reputation, or credibility of governments, so that social frustration and anger is low. In this case, the public simply has no reason to join collective action. This type of censorship can happen in both autocracies and democracies. Finally, censorship isolates dissent groups. If the dissent group wants to expand by seeking for new members, signals have to be delivered to the public. However, censorship can successfully cut off the channel whereby dissents speak to the public. Further, when the state uses selective repression on the dissent group, such as killings and imprisonment, which can impair regime legitimacy, censorship lowers the damage by not letting the public know about the state’s violation of human rights. Hence, the possibility of the aforementioned backfire effect is low. 

\textit{Hypothesis 2: Information repression (censorship) deters mass mobilization.}

\vspace{3mm}

\section{Research Design}

In this research, the unit of analysis is country-year, and the data cover 162 countries between 1990 and 2016. The dependent variables are \textit{the number of total protest} and \textit{the numbers of violent protest}, from @ClarkandRegan2016. There are four independent variables of repression: \textit{physical rights repression}, \textit{civil society repression}, \textit{government’s effort on media censorship}, and \textit{government’s effort on internet censorship}. The first two variables are from CIRI Human Rights project, and the last two variables are from V-Dem version 9. Originally, all these four variables are coded in a way of measuring government’s respect for these human rights, that higher values indicate lower repression. To make the results more intuitive, I recoded these four variables reversely, that higher values indicate higher repression. Control variables include: \textit{GDP per capita (logged)}, \textit{total population (logged)}, and \textit{democracy index}. The first two variables are from World Bank, and the variable of democracy is from V-Dem version 9. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# load all the packages you will use below

library(dotwhisker)
library(tidyverse)
library(haven)
library(plm)
library(dotwhisker)
library(broom)
library(foreign)
library(lme4)
library(stargazer)
library(ggplot2)
library(broom)
library(dplyr)
df<-read_dta("replication_data.dta")
df2<-df %>%
  mutate(physint=8-physint, civil_repression=10-civil_repression, v2mecenefm=-v2mecenefm, v2mecenefi=-v2mecenefi)%>%
  mutate(l_physint=lag(physint, 1), l_civil_repression=lag(civil_repression, 1), l_media=lag(v2mecenefm, 1), l_internet=lag(v2mecenefi, 1))

df_tidy<-pdata.frame(df2, index = c("cowcode","year"))

df_sub1<-df2%>%
  filter(year>=1990) %>%
  select(country, cowcode, year, total_protest, violent_protest,region) %>%
  group_by(region,year)%>%
  summarize(sum.total=sum(total_protest, na.rm = TRUE), sum.violent=sum(violent_protest, na.rm = TRUE)) %>%
  filter(complete.cases(region)) %>%
  mutate(region_factor=recode(region, "1"="Europe", "2"="Middle East", "3"="Asia","4"="Africa","5"="Americas"))

df_sub2<-df2%>%
  filter(year>=1990) %>%
  select(country, cowcode, year, total_protest, violent_protest) %>%
  group_by(year)%>%
  summarize(sum.total=sum(total_protest, na.rm = TRUE), sum.violent=sum(violent_protest, na.rm = TRUE)) %>%
  filter(complete.cases(.)) 
```

\section{Description of Protest}

Graph 1 and Graph 2 show the numbers of total protest and violent protest in the world over time. Both graphs present that, though the number of protest fluctuates over time, the recent decade has witnessed a sharp growth of mass mobilization after 2010. Graph 3 and Graph 4 show the numbers of total protest and violent protest by region. Overall, Middle East has the lowest level of total protest and violent protest, Africa and Europe have higher standard deviation compared with other regions. 

```{r, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE)
##describe data##
graph1<-ggplot(data = df_sub2, aes(year, sum.total))+geom_point()+theme_bw()+ylab("Total Protest")+scale_colour_hue(name = "Region")
graph2<-ggplot(data = df_sub2, aes(year, sum.total))+geom_point()+theme_bw()+ylab("Violent Protest")+scale_colour_hue(name = "Region")

graph3<-ggplot(data = df_sub1, aes(year, sum.total, colour=region_factor))+geom_violin()+theme_bw()+ylab("Total Protest")+scale_colour_hue(name = "Region")
graph4<-ggplot(data = df_sub1, aes(year, sum.violent, colour=region_factor))+geom_violin()+theme_bw()+ylab("Violent Protest")+scale_colour_hue(name = "Region")
```

\begin{center}
    \textbf{Graph1}
\end{center}

```{r, echo=FALSE}
graph1
```

\begin{center}
    \textbf{Graph2}
\end{center}

```{r, echo=FALSE}
graph2
```

\begin{center}
    \textbf{Graph3}
\end{center}

```{r, echo=FALSE}
graph3
```

\begin{center}
    \textbf{Graph4}
\end{center}

```{r, echo=FALSE}
graph4
```

\section{Empirical Evidence}

I run a panel regression model with random effects to see how different repression strategies shape collective action behavior. The results are shown in Graph 5 and Graph 6. The dependent variable in Graph 5 is \textit{total protest}. The orange line is the baseline model, in where only four independent variables of repression strategies are included. The blue line represents the full model, where all control variables are included and region is controlled for. Graph 5 shows that \textit{physical rights repression} is positively correlated with \textit{total protest} while \textit{media censorship} is negatively correlated with \textit{total protest}, and these two results are significant in both models. That is to say, physical rights repression tends to escalate mass mobilization, while media censorship can effectively deter it. Graph 6 is similar, except that the dependent variable is \textit{violent protest}. Likewise, physical rights repression escalates violent mass mobilization, and media censorship deters it. Both hypothesis 1 and hypothesis 2 are confirmed. Another thing to be noticed is that, civil rights repression shows insignificant results in three out of four models. It is only significant when the dependent variable is violent protest and other control variables are added in. This result indicates that the impact of civil rights repression is not robust, that it can deter mobilization in some situation, but stays ineffective in some other situation. 

```{r, echo=FALSE}
##total protest as DV##
model1<-plm(total_protest~l_physint+l_civil_repression+l_media+l_internet, data = df_tidy, model = "random")
model2<-plm(total_protest~l_physint+l_civil_repression+l_media+l_internet+lngdp+lnpop+v2x_polyarchy+Europe+Middle_East+Asia+Africa, data = df_tidy, model = "random")

graph5<-dwplot(list(model1, model2)) %>%
  relabel_predictors(c(l_physint ="Physical Rights Repression", l_civil_repression ="Civil Rights Repression", l_media="Media Censorship", l_internet="Internet Censorship", lngdp="GDP per capita", lnpop="Total Population",v2x_polyarchy="Democracy", Middle_East="Middle East" ))+theme_bw()+xlab("Coefficient")+geom_vline(xintercept = 0, colour="grey60", linetype=2)+scale_colour_hue(name = "Model", labels = c("Baseline Model", "Full Model"))+ggtitle("Repression and Total Protest")

##Violent protest as DV##
model3<-plm(violent_protest~l_physint+l_civil_repression+l_media+l_internet, data = df_tidy, model = "random")
model4<-plm(violent_protest~l_physint+l_civil_repression+l_media+l_internet+lngdp+lnpop+v2x_polyarchy+Europe+Middle_East+Asia+Africa, data = df_tidy, model = "random")

graph6<-dwplot(list(model3,model4))%>%
  relabel_predictors(c(l_physint ="Physical Rights Repression", l_civil_repression ="Civil Rights Repression", l_media="Media Censorship", l_internet="Internet Censorship", lngdp="GDP per capita", lnpop="Total Population",v2x_polyarchy="Democracy", Middle_East="Middle East" ))+theme_bw()+xlab("Coefficient")+geom_vline(xintercept = 0, colour="grey60", linetype=2)+scale_colour_hue(name = "Model", labels = c("Baseline Model", "Full Model"))+ggtitle("Repression and Violent Protest")
```

\begin{center}
    \textbf{Graph5}
\end{center}

```{r, echo=FALSE}
graph5
```

\begin{center}
    \textbf{Graph6}
\end{center}

```{r, echo=FALSE}
graph6
```

\section{Conclusion}

Previous studies have provided competing theories and mixed empirical evidence on how repression effects mass mobilization. This paper shows that, the answer to this question lies in different repression strategies. Physical rights repression, such as killing, imprisonment, and torture, can damage regime legitimacy and even motivates the public to join dissent groups. Hence, physical rights repression tends to have a backfire effect that it escalates mass mobilization. In contrast, censorship can effectively deter mass mobilization by isolating the country from outside world, isolating the dissent group from the public, and eliminating negative news on government. Thus, censorship lowers the willingness for the public to take part in collective action and also reduces the opportunities for the dissent group to expand. Hence, censorship is in fact the most effective repression strategy. 

\section{Appendix}
```{r, echo=FALSE, include=FALSE}
table1<-stargazer(model1, model2, model3, model4)
```
\begin{table}[!htbp] \centering 
  \caption{Panel Regression with Random Effect} 
  \label{} 
\begin{tabular}{@{\extracolsep{5pt}}lcccc} 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & \multicolumn{4}{c}{\textit{Dependent variable:}} \\ 
\cline{2-5} 
\\[-1.8ex] & \multicolumn{2}{c}{Total Protest} & \multicolumn{2}{c}{Violent Protest} \\ 
\\[-1.8ex] & (1) & (2) & (3) & (4)\\ 
\hline \\[-1.8ex] 
 Physical Rights Repression & 0.360$^{***}$ & 0.298$^{***}$ & 0.149$^{***}$ & 0.096$^{***}$ \\ 
  & (0.065) & (0.070) & (0.023) & (0.025) \\ 
  & & & & \\ 
 Civil Rights Repression & $-$0.058 & $-$0.113$^{*}$ & $-$0.019 & $-$0.053$^{**}$ \\ 
  & (0.065) & (0.068) & (0.023) & (0.025) \\ 
  & & & & \\ 
 Media Censorship & $-$0.528$^{***}$ & $-$0.514$^{**}$ & $-$0.229$^{***}$ & $-$0.359$^{***}$ \\ 
  & (0.162) & (0.199) & (0.056) & (0.070) \\ 
  & & & & \\ 
 Internet Censorship & 0.104 & 0.106 & 0.071 & 0.023 \\ 
  & (0.172) & (0.185) & (0.058) & (0.063) \\ 
  & & & & \\ 
 GDP per capita &  & 0.293 &  & $-$0.019 \\ 
  &  & (0.186) &  & (0.060) \\ 
  & & & & \\ 
 Total Population &  & 0.684$^{***}$ &  & 0.241$^{***}$ \\ 
  &  & (0.156) &  & (0.049) \\ 
  & & & & \\ 
 Democracy &  & $-$1.867 &  & $-$1.556$^{***}$ \\ 
  &  & (1.281) &  & (0.449) \\ 
  & & & & \\ 
 Europe &  & 0.252 &  & $-$0.118 \\ 
  &  & (0.721) &  & (0.223) \\ 
  & & & & \\ 
 Middle East &  & $-$0.830 &  & 0.017 \\ 
  &  & (1.053) &  & (0.330) \\ 
  & & & & \\ 
 Asia &  & $-$0.014 &  & 0.012 \\ 
  &  & (0.814) &  & (0.253) \\ 
  & & & & \\ 
 Africa &  & $-$0.835 &  & $-$0.058 \\ 
  &  & (0.735) &  & (0.229) \\ 
  & & & & \\ 
 Constant & 1.934$^{***}$ & $-$9.804$^{***}$ & 0.196 & $-$2.473$^{***}$ \\ 
  & (0.415) & (2.907) & (0.142) & (0.914) \\ 
  & & & & \\ 
\hline \\[-1.8ex] 
Observations & 2,628 & 2,580 & 2,628 & 2,580 \\ 
R$^{2}$ & 0.013 & 0.026 & 0.019 & 0.035 \\ 
Adjusted R$^{2}$ & 0.012 & 0.022 & 0.018 & 0.031 \\ 
F Statistic & 35.406$^{***}$ & 69.547$^{***}$ & 51.966$^{***}$ & 92.826$^{***}$ \\ 
\hline 
\hline \\[-1.8ex] 
\textit{Note:}  & \multicolumn{4}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\ 
\end{tabular} 
\end{table}


